---
title: "California 2022 Proposition 30"
author: "GitData"
date: "10/19/2022"
output:
  html_document:
    toc: yes
    toc_depth: '5'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 5
---

\clearpage
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(knitr)
library(ggplot2)
library(dplyr)
library(GGally)
library(lmtest)
library(car)
library(onewaytests)
require(ggiraph)
require(ggiraphExtra)
require(plyr)
library(ggpubr)
```

# I. Introduction and Research Questions

California has been a pioneer in U.S. air quality legislation since the 1940s. In 2006 California passed Assembly Bill 32 (AB 32), which created a comprehensive, multi-year program to reduce greenhouse gas (GHG) emissions in California. In 2050, the emission of greenhouse gas should be cut to 80% below the 1990 level. According to the California Air Resource Board, transportation produced about 41% of greenhouse gas emissions in 2019. Moreover, the California Air Resource Board reported that the wildfire produced 106.7 million tons of carbon dioxide in 2020 and 85.1 million tons in 2021. The state wants to decrease greenhouse gas emissions by having more electric cars and preventing wildfires. This is what Proposition 30 is going to do.

Proposition 30 is going to increase a 1.75% tax on Californians who earn above $2 million per year to fund programs to reduce greenhouse gas emissions and prevent wildfires. The tax plans to start in 2023 and will end by 2043 or earlier if we can maintain the expected amount of greenhouse gas emissions for three consecutive calendar years.

80 percent of the tax revenue will be invested in the Zero Emission Vehicle (ZEV) program. Specifically, 45 percent in helping people buy electric vehicles and 35 percent in installing more charging stations. The state government decided to dedicate an additional $10 billion over the next five years to the program in this year’s budget.

20 percent of the tax revenue will be spent on Wildfire Response and Prevention Programs. In general, most of this money will be used to hire, train, and retain state firefighters. The rest of the money could be used for other wildfire response and prevention activities.

We will figure out whether more electric vehicles can help reduce greenhouse gas emissions. Based on the tax revenue and the additional $10 billion funded by the government, we will find out how many electric vehicles can be used and the amount of greenhouse gas that can be reduced. We will find out how much greenhouse gas can be reduced after funding the Wildfire Response and Prevention Programs. Finally, we will evaluate if our goal can be achieved, which is whether greenhouse gas emissions can be cut to 80% below the 1990 level by 2050.


# II. Data Information
```{r, include=FALSE}
################################################
##########II. Data Information##################
################################################
# Load dataset
pollution = read.csv("https://caelection2022.gitdata.ga/dataset/EMFAC2021-EI-202x.csv", 
                     header = TRUE)
income = read.csv("https://caelection2022.gitdata.ga/dataset/2020_Combined_AR_PIT_B4A.csv", 
                    header = TRUE)
fuelStation = read.csv("https://caelection2022.gitdata.ga/dataset/alt_fuel_stations.csv",
                       header = TRUE)
carSale = read.csv("https://caelection2022.gitdata.ga/dataset/New_ZEV_Sales_County.csv", 
                    header = TRUE)
# Pollution per mile per car
pollutionAVG = pollution %>% 
                         mutate(CO2_AVG = CO2_RUNEX / Total.VMT / 1000000,
                                CH4_AVG = CH4_RUNEX / Total.VMT / 1000000,
                                N2O_AVG = N2O_RUNEX / Total.VMT / 1000000,
                                GHG_AVG = (CO2_RUNEX + CH4_RUNEX + N2O_RUNEX) / Total.VMT / 1000000,
                                CAR_AGE = case_when(Model.Year >= 1996 ~ "New", 
                                          Model.Year < 1996 ~ "Old"))
```

Before figuring out whether the fuel type of the cars affect the emission of greenhouse gas, including Carbon dioxide, Methane, and Nitrous oxide, we need to know some basic aspects of the data we have.

- `Fuel`: The types of fuel each car uses. There are four types of fuel: `Diesel` (37,968 counts), `Electricity` (27,314 counts), `Gasoline` (40,140 counts), and `Plug-in Hybrid` (11,604 counts).

- `CO2_AVG`: The amount of Carbon dioxide, in million metric tons, emitted per mile per car in 2021.

- `CH4_AVG`: The amount of Methane, in million metric tons, emitted per mile per car in 2021.

- `N2O_AVG`: The amount of Nitrous oxide, in million metric tons, emitted per mile per car in 2021.

- `GHG_AVG`: The amount of greenhouse gas (the sum of Carbon dioxide, Methane, and Nitrous oxide), in million metric tons, emitted per mile per car in 2021.


# III. Data Visualization

```{r}
boxplot(pollutionAVG$CO2_AVG ~ pollutionAVG$Fuel)
co2_bar = ggplot(pollutionAVG, aes(y = CO2_AVG, x = Fuel)) +
  geom_bar(position='dodge', stat='identity')
ch4_bar = ggplot(pollutionAVG, aes(y = CH4_AVG, x = Fuel)) +
  geom_bar(position='dodge', stat='identity')
n2o_bar = ggplot(pollutionAVG, aes(y = N2O_AVG, x = Fuel)) +
  geom_bar(position='dodge', stat='identity')
ghg_bar = ggplot(pollutionAVG, aes(y = GHG_AVG, x = Fuel)) +
  geom_bar(position='dodge', stat='identity')
ggarrange(co2_bar, ch4_bar, n2o_bar, ghg_bar,  
          labels = c("A", "B", "C", "D"),
          ncol = 2, nrow = 2)
```

```{r}
################################################
##########III. Data Visualization###############
################################################
# Linear regression for CO2
co2.model.avg = lm(CO2_AVG ~ as.character(Fuel), data = pollutionAVG)
summary(co2.model.avg)
#co2.model.avg = lm(CO2_AVG ~ as.character(Fuel) + as.character(CAR_AGE) + 
#                             Speed, data = pollutionAVG)
#summary(co2.model.avg)
```

```{r}
# Linear regression for CH4
ch4.model.avg = lm(CH4_AVG ~ as.character(Fuel), data = pollutionAVG)
summary(ch4.model.avg)
#ch4.model.avg = lm(CH4_AVG ~ as.character(Fuel) + as.character(CAR_AGE) + 
#                             Speed, data = pollutionAVG)
#summary(ch4.model.avg)
```

```{r}
# Linear regression for N2O
n2o.model.avg = lm(N2O_AVG ~ as.character(Fuel), data = pollutionAVG)
summary(n2o.model.avg)
#n2o.model.avg = lm(N2O_AVG ~ as.character(Fuel) + as.character(CAR_AGE) + 
#                             Speed, data = pollutionAVG)
#summary(n2o.model.avg)
```

```{r}
# Linear regression for GHG
ghg.model.avg = lm(GHG_AVG ~ as.character(Fuel), data = pollutionAVG)
summary(ghg.model.avg)
#ghg.model.avg = lm(GHG_AVG ~ as.character(Fuel) + as.character(CAR_AGE) + 
#                             Speed, data = pollutionAVG)
#summary(ghg.model.avg)
```

```{r}
# Tax
num_people = 14911 + 6280 + 3479 + 10335
num_people
increase_tax = (34229385 + 20578127 + 14510953 + 150335815) * 0.0175 * 1000
increase_tax
buy_car = increase_tax * 0.45
charging_station = increase_tax * 0.35
fire_control = increase_tax * 0.2
# Only for car and charging station in the next 5 years
addition = 10000000000 
addition_car_year = addition / 5 * 0.5625
addition_charging_year = addition / 5 * 0.4375
total_car_2027_year = buy_car + addition_car_year
total_charging_2027_year = charging_station + addition_charging_year
```


# IV. Analysis

// Type your solution here

# V. Result and Interpretation

// Type your solution here

# VI. Conclusion and Future Discussion

// Type your solution here


\clearpage

# Appendix: R Script

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```